---
title: "Campaign Finance Exploration"
author: "Chris Hua"
date: "17 December 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(magrittr)
library(ggplot2)
DATA_SRC <- "https://data.world/codefordc/campaign-finance"
saved_cfg <- data.world::save_config("asdf")
data.world::set_config(saved_cfg)

```

## Motivation

Campaign finance is a well studied and deeply interesting area of politics. This writeup explores some campaign donations from Washington DC.

## The Data

Just a few rows

```{r}
qry <- data.world::qry_sql(
    "SELECT * FROM contributions LIMIT 5"
)
data.world::query(qry, DATA_SRC) %>% knitr::kable()
```

### Number of donations by time

Next, we want to see if there's any interesting trends in the donation dates and aggregated amounts. We can pull this data with the following SQL code:

```{r}
date_qry <- data.world::qry_sql("SELECT 
    date_of_receipt     AS dt
    , COUNT(*)          AS n_donations
    , SUM(amount)       AS sum_donations
    FROM contributions
    GROUP BY dt
    ORDER BY dt")
date_df <- data.world::query(date_qry, DATA_SRC)
```

At a date granularity, we can kind of see some trends in the data, but it's really noisy. 

We can plot the donations, with one point per day.

```{r}
date_df %>%
    ggplot(aes(x = dt, y = n_donations)) +
    geom_point() + 
    ggtitle("Donations by date") + 
    ylab("# of Donations") + 
    xlab("Date")+
    scale_x_date(date_breaks = "1 years", date_labels = "%Y")
```

Let's group the donations by week instead. This looks much cleaner!

```{r}
date_df %>%
    group_by(dt = lubridate::round_date(dt, unit = "week")) %>%
    summarize(n_donations = sum(n_donations),
              amount = sum(sum_donations)) %>%
    ggplot(aes(x = dt, y = n_donations)) + 
    geom_point() +
    ggtitle("Donations by date", "Rounded to weekly data") + 
    ylab("# of Donations") + 
    xlab("Date")+
    scale_x_date(date_breaks = "1 years", date_labels = "%Y")
```

Finally, let's group them by month and visualize the results. When working with data, we often have to trade off human interpretability for robustness. 

```{r}
date_df %>%
    group_by(dt = lubridate::round_date(dt, unit = "month")) %>%
    summarize(n_donations = sum(n_donations),
              amount = sum(sum_donations)) %>%
    ggplot(aes(x = dt, y = n_donations)) + 
    geom_point() +
    ggtitle("Donations by date", "Rounded to monthly data") + 
    ylab("# of Donations") + 
    xlab("Date") +
    scale_x_date(date_breaks = "1 years", date_labels = "%Y")
```

### Amount of donations by time

```{r}
date_df %>%
    group_by(dt = lubridate::round_date(dt, unit = "week")) %>%
    summarize(n_donations = sum(n_donations),
              amount = sum(sum_donations)) %>%
    ggplot(aes(x = dt, y = amount)) + 
    geom_point() +
    ggtitle("Sum of donations by date", "Rounded to weekly data") + 
    ylab("Amount USD") + 
    xlab("Date")+
    scale_x_date(date_breaks = "1 years", date_labels = "%Y")
```

```{r}
date_df %>%
    group_by(dt = lubridate::round_date(dt, unit = "month")) %>%
    summarize(n_donations = sum(n_donations),
              amount = sum(sum_donations)) %>%
    ggplot(aes(x = dt, y = amount)) + 
    geom_point() +
    ggtitle("Sum of donations by date", "Rounded to weekly data") + 
    ylab("Amount USD") + 
    xlab("Date")+
    scale_x_date(date_breaks = "1 years", date_labels = "%Y")
```

### Campaign level data

We can easily pull some campaign level data to figure out which campaigns raised the most money and garnered the most engagement.

```{r}
cmp_qry <- data.world::qry_sql("SELECT 
    committee_name      AS campaign
    , date_of_receipt   AS dt
    , COUNT(*)          AS n_donations
    , SUM(amount)       AS sum_donations
    FROM contributions
    GROUP BY campaign, dt
    ORDER BY campaign, dt")
cmp_df <- data.world::query(cmp_qry, DATA_SRC)
```



```{r}
cmp_df %>%
    group_by(campaign, dt = lubridate::round_date(dt, "year")) %>%
    summarize(n_donations = sum(n_donations), 
              sum_donations = sum(sum_donations)) %>%
    arrange(-n_donations) %>%
    head(10)
```

```{r}
cmp_df %>%
    group_by(campaign, dt = lubridate::round_date(dt, "year")) %>%
    summarize(n_donations = sum(n_donations), 
              sum_donations = sum(sum_donations)) %>%
    arrange(-sum_donations) %>% 
    head(10)
```
