---
title: "Bitcoin Hashing Example"
author: "Chris Hua"
date: "12/25/2017"
output: 
  html_document: 
    keep_md: yes
    self_contained: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F)
library(digest)
library(dplyr)
library(magrittr)
```

# Part 1: Create Keys

As per class, we define some $n = p*q$, or `r 7 * 23`. Then, given $e=13$, the public key is a composition of 161 and 13. 

We can find the private key by evaluating the modulo condition for each $d$ between 1 and a given max number, then returning the first one that is equal to 1. This operation is vectorized and as such much faster in R than a loop. 

```{r, echo= T}
find_priv_key <- function(p, q, e, max_iter = 100) {
    d <- 1:max_iter
    val <- (e * d) %% ((p - 1) * (q - 1))
    min(which(val == 1))
}
find_priv_key(7, 23, 13)
```

The private key is then 61. 

## Part 2: Sign a message

To do the modulo math necessary to sign a message, we need to do modular exponentation (generally). We borrow the implementation in package `numbers` to do this.

```{r, echo= T}
sign_message <- function(pk, n, m) {
    numbers::modpower(m, pk, n)
}
sign_message(61.0, 161.0, 12.0)
```

## Part 3: Check if a message is authentic

To check if a message is authentic we can create another function to check the encryption.

```{r}
check_message <- function(m, c, e, pk) {
    numbers::modpower(c, e, pk) == m
}

# public key stuff
pk = 161
e = 13

# example from slides works...
# check_message(2, 8, 7, 55)
```

Message 1: 

```{r}
c_1 = c(210, 453)
m_1 = c(10, 24)
check_message(m_1, c_1, e, pk)
```

Message 2:

```{r}
c_2 = c(519, 370)
m_2 = c(11, 30)
check_message(m_2, c_2, e, pk)
```

Message 3:

```{r}
c_3 = c(12, 594)
m_3 = c(12, 16)
check_message(m_3, c_3, e, pk)
```

None of these are authentic!

# Part 4: Mining Bitcoin

We find a nonce that, when appended to the message and hashed via md5, starts with 3 zeroes. Our process is as follows:

```{r, echo=T}
prove_work <- function(m, k, start = 1000, max_iter = 5000) {
    for(i in start:(start+max_iter)) {
        item <- paste0(m, i)
        hash <- digest::digest(item, algo = "md5")
        if(substr(hash, start = 1, stop = 3) == "000") {
            return(i)
        }
    }
    -1 # answer is outside search bounds
}
```

Then our result is given by:

```{r}
prove_work(2017, 3)
```

# Part 5: Bitcoin game theory

**a)** There is some optimum search start point, we can call it $i^{\prime}$. Since we have approximately equivalent compute power, we will assume that we can check each number in equal amounts of time.

Then, if my competition starts at 0, and I start at 1, in expectation I will dominate them, since I will check each number greater than 1, *before* they can. 

We don't show it here but the greatest concentration of "hash solutions" is at a relatively low number. We performed a search for 1000 hash solutions with $k=3$ digits, and with hash values between 0 and 20000, and examined the distribution. The average value was about 3600, and 80% were under 5000. The insight is that there is a long tail of values to check for any given message, and that there might be increased efficiency if we do a greedy search - cap the number of iterations we try for each block, and aim to solve as many 'easy' blocks as possible.

We should combine the 'greedy' approach with a larger start value in our search against otherwise clueless competitors.

**b)** In the face of competition using the same strategy, we would move to larger start numbers. As we noted above, there is a tradeoff between the start point $i$ and the number of blocks solvable within $k$ iterations from the start point.


